---
title: T-Rex Selector Algorithm
output:
  md_document:
    variant: gfm
  html_document:
    keep_md: true
  pdf_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
# Store user's options()
old_options <- options()

if (!dir.exists("man/figures/")) {
  dir.create("man/figures/")
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TRexSelectorNeo

**Title**: The T-Rex selector for fast high-dimensional variable selection with FDR control

**Description**: It performs fast variable selection in large-scale high-dimensional settings while controlling the false discovery rate (FDR) at a user-defined target level. The R package provides bindings to the TRexSelector C++ library and supersedes the earlier pure-R 'TRexSelector' package, alongside which it can be installed.

**Papers**: The package is based on the papers

J. Machkour, M. Muma, and D. P. Palomar, "The terminating-random experiments selector: Fast high-dimensional variable selection with false discovery rate control," *Signal Processing*, vol. 231, article 109894, 2025. (<https://doi.org/10.1016/j.sigpro.2025.109894>)

J. Machkour, M. Muma, and D. P. Palomar, "High-dimensional false discovery rate control for dependent variables," *Signal Processing*, vol. 234, article 109990, 2025. (<https://doi.org/10.1016/j.sigpro.2025.109990>)

**Note**: The T-Rex selector performs terminated-random experiments (T-Rex) using the T-LARS algorithm and fuses the selected active sets of all random experiments to obtain a final set of selected variables. The T-Rex selector provably controls the false discovery rate (FDR), i.e., the expected fraction of selected false positives among all selected variables, at the user-defined target level while maximizing the number of selected variables and, thereby, achieving a high true positive rate (TPR) (i.e., power). The T-Rex selector can be applied in various fields, such as genomics, financial engineering, or any other field that requires a fast and FDR-controlling variable/feature selection method for large-scale high-dimensional settings.

In the following sections, we show you how to install and use the package.

## Installation

The package compiles its C++ backend on installation, so a C++20-capable compiler is required (GCC ≥ 10, Clang ≥ 12, Apple Clang ≥ 14, or Rtools on Windows), together with R ≥ 4.3.

First install the R-side prerequisites (once, from an R session):

```{r, eval=FALSE}
install.packages(c("Rcpp", "RcppEigen", "BH", "Rcereal", "R6"))
```

Then install 'TRexSelectorNeo' from [GitHub](https://github.com/G4Lactus/TRexSelector) — the package lives in the `R/TRexSelectorNeo/` subdirectory of the repository:

```{r, eval=FALSE}
# Option 1: Install directly from GitHub
install.packages("remotes")
remotes::install_github("G4Lactus/TRexSelector", subdir = "R/TRexSelectorNeo")

# Option 2: Install from a local clone of the repository
# (from the shell, at the repository root)
# R CMD INSTALL R/TRexSelectorNeo
```

*(A CRAN submission is in progress; once accepted, `install.packages("TRexSelectorNeo")` will provide the stable version.)*

You can open the help pages with:
```{r, eval=FALSE}
library(TRexSelectorNeo)
help(package = "TRexSelectorNeo")
?TRexSelector
?trex_control
?compute_fdp
?compute_tpp
# etc.
```

To cite the package 'TRexSelectorNeo' in publications use:
```{r, eval=FALSE}
citation("TRexSelectorNeo")
```

# Quick Start
This section illustrates the basic usage of the 'TRexSelectorNeo' package to perform FDR-controlled variable selection in large-scale high-dimensional settings based on the T-Rex selector.

1. **First**, we generate a high-dimensional Gaussian data set with sparse support:

```{r}
library(TRexSelectorNeo)

# Setup
n <- 75 # number of observations
p <- 150 # number of variables
num_act <- 3 # number of true active variables
beta <- c(rep(1, times = num_act), rep(0, times = p - num_act)) # coefficient vector
true_actives <- which(beta > 0) # indices of true active variables

# Generate Gaussian data
set.seed(123)
X <- matrix(stats::rnorm(n * p), nrow = n, ncol = p)
y <- X %*% beta + stats::rnorm(n)
```

2. **Second**, we perform FDR-controlled variable selection using the T-Rex selector for a target FDR of 5%. The selectors are R6 classes: construct with `$new(...)`, run with `$select()`, and read the results from active fields such as `$selected_indices` (1-based indices):

```{r}
# Seed
set.seed(1234)

# Variable selection via T-Rex
sel <- TRexSelector$new(X, y, tFDR = 0.05, verbose = FALSE)
sel$select()

paste0("True active variables: ", paste(as.character(true_actives), collapse = ", "))
paste0("Selected variables:    ", paste(as.character(sel$selected_indices), collapse = ", "))
```

So, for a preset target FDR of 5%, the T-Rex selector has selected all true active variables and there are no false positives in this example.

Note that users have to choose the target FDR according to the requirements of their specific applications.

All selector variants (`TRexSelector`, `TRexDASelector`, `TRexGVSSelector`, `TRexScreeningSelector`, `TRexBiobankScreeningSelector`, `TRexSPCASelector`) follow the same construct → `$select()` → inspect pattern.

## Documentation
For more information and some examples, please check the package vignette in [`vignettes/`](vignettes/) (built on installation with `build_vignettes = TRUE`).

## Links
T-Rex paper: <https://doi.org/10.48550/arXiv.2110.06048>

TRexSelector repository (C++ core, Python, and R packages): [GitHub-TRexSelector](https://github.com/G4Lactus/TRexSelector)

Bug reports: [GitHub Issues](https://github.com/G4Lactus/TRexSelector/issues)

```{r, include = FALSE}
# Reset user's options()
options(old_options)
```
