---
title: T-Rex Selector Algorithm
output:
  md_document:
    variant: gfm
  html_document:
    keep_md: true
  pdf_document: default
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
# Store user's options()
old_options <- options()

if (!dir.exists("man/figures/")) {
  dir.create("man/figures/")
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TRexSelector

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/TRexSelector)](https://CRAN.R-project.org/package=TRexSelector)
[![CRAN Downloads](https://cranlogs.r-pkg.org/badges/TRexSelector)](https://CRAN.R-project.org/package=TRexSelector)
![CRAN Downloads Total](https://cranlogs.r-pkg.org/badges/grand-total/TRexSelector?color=brightgreen)

**Title**: The T-Rex selector for fast high-dimensional variable selection with FDR control

**Description**: It performs fast variable selection in large-scale high-dimensional settings while controlling the false discovery rate (FDR) at a user-defined target level. The R package provides bindings to the TRexSelector C++ library.

**Papers**: The package is based on the papers

J. Machkour, M. Muma, and D. P. Palomar, "The terminating-random experiments selector: Fast high-dimensional variable selection with false discovery rate control," *Signal Processing*, vol. 231, article 109894, 2025. (<https://doi.org/10.1016/j.sigpro.2025.109894>)

J. Machkour, M. Muma, and D. P. Palomar, "High-dimensional false discovery rate control for dependent variables," *Signal Processing*, vol. 234, article 109990, 2025. (<https://doi.org/10.1016/j.sigpro.2025.109990>)

**Note**: The T-Rex selector performs terminated-random experiments (T-Rex) using the T-LARS algorithm and fuses the selected active sets of all random experiments to obtain a final set of selected variables. The T-Rex selector provably controls the false discovery rate (FDR), i.e., the expected fraction of selected false positives among all selected variables, at the user-defined target level while maximizing the number of selected variables and, thereby, achieving a high true positive rate (TPR) (i.e., power). The T-Rex selector can be applied in various fields, such as genomics, financial engineering, or any other field that requires a fast and FDR-controlling variable/feature selection method for large-scale high-dimensional settings.

In the following sections, we show you how to install and use the package.

## Installation

You can install the 'TRexSelector' package from [CRAN](https://CRAN.R-project.org/package=TRexSelector) (stable version) or [GitHub](https://github.com/G4Lactus/TRexSelector) (developer version) with:

```{r, eval=FALSE}
# Option 1: Install stable version from CRAN
install.packages("TRexSelector")

# Option 2: Install developer version from GitHub
install.packages("devtools")
devtools::install_github("G4Lactus/TRexSelector")
```

You can open the help pages with:
```{r, eval=FALSE}
library(TRexSelector)
help(package = "TRexSelector")
?TRexSelector
?trex_control
?compute_fdp
?compute_tpp
# etc.
```

To cite the package 'TRexSelector' in publications use:
```{r, eval=FALSE}
citation("TRexSelector")
```

# Quick Start
This section illustrates the basic usage of the 'TRexSelector' package to perform FDR-controlled variable selection in large-scale high-dimensional settings based on the T-Rex selector.

1. **First**, we generate a high-dimensional Gaussian data set with sparse support:

```{r}
library(TRexSelector)

# Setup
n <- 75 # number of observations
p <- 150 # number of variables
num_act <- 3 # number of true active variables
beta <- c(rep(1, times = num_act), rep(0, times = p - num_act)) # coefficient vector
true_actives <- which(beta > 0) # indices of true active variables

# Generate Gaussian data
set.seed(123)
X <- matrix(stats::rnorm(n * p), nrow = n, ncol = p)
y <- X %*% beta + stats::rnorm(n)
```

2. **Second**, we perform FDR-controlled variable selection using the T-Rex selector for a target FDR of 5%:

```{r}
# Seed
set.seed(1234)

# Variable selection via T-Rex
sel <- TRexSelector$new(X, y, tFDR = 0.05, verbose = FALSE)
sel$select()

paste0("True active variables: ", paste(as.character(true_actives), collapse = ", "))
paste0("Selected variables:    ", paste(as.character(sel$selected_indices), collapse = ", "))
```

So, for a preset target FDR of 5%, the T-Rex selector has selected all true active variables and there are no false positives in this example.

Note that users have to choose the target FDR according to the requirements of their specific applications.

## Documentation
For more information and some examples, please check the [GitHub-vignette](https://htmlpreview.github.io/?https://github.com/G4Lactus/TRexSelector/blob/main/vignettes/TRexSelector_usage_and_simulations.html).

## Links
T-Rex paper: https://doi.org/10.48550/arXiv.2110.06048

TRexSelector package (stable version): [CRAN-TRexSelector](https://CRAN.R-project.org/package=TRexSelector).

TRexSelector package (developer version): [GitHub-TRexSelector](https://github.com/G4Lactus/TRexSelector).

README file: [GitHub-readme](https://htmlpreview.github.io/?https://github.com/G4Lactus/TRexSelector/blob/main/README.html).

Vignette: [GitHub-vignette](https://htmlpreview.github.io/?https://github.com/G4Lactus/TRexSelector/blob/main/vignettes/TRexSelector_usage_and_simulations.html).

```{r, include = FALSE}
# Reset user's options()
options(old_options)
```
